### 2017-04-22 paper.tex
Started on rough rough draft

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{color}

\begin{document}
\section*{Modified Teaching environment}
We will solve a more relaxed problem such that the teacher may not only give $(x,y)$ training examples but may also give single $x$ values to probe the learner for output.

\section*{Theorem 1}
If given a perceptron with known starting parameter $w_0$, desired parameter $w^*$, and unknown learning rate $\eta \in [\alpha , \beta]$ for given $\epsilon > 0$, we can guarantee $\eta \in [\alpha', \beta']$ such that $\beta' - \alpha' \leq \epsilon$ using $1 + \lceil \log_2(\frac{\beta - \alpha}{\epsilon})\rceil$

\subsection*{Proof:}
Assume $w_0 \neq cw^*$ for arbitrary $c > 0$, then there exists an $(x_1, y_1)$ such that $\text{sign}(\langle w^*,x_1 \rangle) = y_1 = 1$ and $\text{sign}(\langle w_0,x_1 \rangle) = -y_1$.
\color{red}
I don't think that its hard to show this is true.
\color{black}

Then give $(x_1, y_1)$ to the learner and then after training update,
$$
w_1 = w_0 + \eta* x_1
$$

Then grab an unlabeled $x_2$ such that
$\langle w_0,x_2 \rangle \neq 0$, $\langle x_1, x_2 \rangle \neq 0$
and 
$$
\langle w_0,x_2 \rangle + \frac{\beta - \alpha}{2} \langle x_1, x_2 \rangle = 0
$$
\color{red}
I'm a moron. What's the closed form of this? I know there is not a unique solution. We just need the simplest one.
\color{black}

Then probe the learner with $x_2$. If the learner outputs $+1$ then
$\eta \in (\frac{\beta - \alpha}{2} , \beta]$, else 
$\eta \in [\alpha,\frac{\beta - \alpha}{2}]$

As you can see our interval for $\eta$ halves with each probe. One can then see that we can iteratively repeat this for a total of $ \lceil \log_2(\frac{\beta - \alpha}{\epsilon})\rceil$ iterations to aquire the desired interval length on $\eta$.

\section*{Theorem 2}
\color{red}

Something to the respect of, if we know $\eta \in [\alpha, \beta]$ then with one training example we can guarantee that
$$
\frac{\langle w^*, w_0 \rangle}{\lvert w^* \rvert \lvert w_0 \rvert} \geq 1 - \epsilon
$$

So by combining with Theorem 1
$$
T_\epsilon \leq \{\text{number of points needed to guarantee } \eta \in [\alpha, \beta]\} +1 
$$
Where $T_\epsilon$ is the $\epsilon$ teaching dimension of the perceptron.
\color{black}

\subsection*{Proof}

\color{red}
Assume we now $w_1$
and $\eta_{min} \leq \eta \leq \eta_{max}$

and there is an $(x_2,y_2)$ such that $y_2 = 1$ and agrees with $w^*$ and not $w_1$. For some $c$ similar to orignal proof
let 
$$
x_2 = \frac{cw^* - w_1}{\beta}
$$

note 
$$
\frac{\alpha}{\alpha + \beta - \alpha}\leq \frac{\eta}{\beta}  \leq 1
$$

so
$$
w_1 + \eta x_2 = cw^* - \epsilon
$$
where
$
0 \leq \epsilon \leq (cw* - w_1) * \frac{\beta - \alpha}{\alpha + \beta - \alpha}  \leq \frac{\beta - \alpha}{\eta_{min} + \beta - \alpha} \leq \frac{\beta - \alpha}{\eta_{min}}
$

\color{red}

\end{document}